---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fbi: R Wrapper for the FBI Crime Data Explorer (CDE) API

<!-- badges: start -->
[![R-CMD-check](https://github.com/Civilytics/fbi_cde/actions/workflows/R-CMD-check.yml/badge.svg)](https://github.com/Civilytics/fbi_cde/actions/workflows/R-CMD-check.yml)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
<!-- badges: end -->

An R wrapper for the [FBI Crime Data Explorer (CDE) API](https://cde.ucr.cjis.gov/LATEST/) providing access to UCR and NIBRS crime data. The package offers a unified interface for querying summarized crime statistics, arrest demographics, NIBRS incident-level data, Supplemental Homicide Reports (SHR), and police employment statistics.

All network requests go through `cde_request()`, the single network seam for the package. **No API key is required** for the current CDE host at `cde.ucr.cjis.gov`.

## Installation

The package is not on CRAN. Install the development version from GitHub:

```r
remotes::install_github("Civilytics/fbi_cde")
```

## Quick Start

```{r example}
library(fbi)

# Get all states available in the API
get_states()

# Get offense codes
get_offense_codes()

# Get summarized crime data for a specific agency
get_agency_crime("CA0010900")

# Get arrest demographics for a state
get_arrest_demographics(state_abb = "CA", offense = "murder")
```

## Table of Contents

- [Lookups](#lookups)
- [Summarized Crime](#summarized-crime)
- [Arrest](#arrest)
- [NIBRS](#nibrs)
- [SHR](#shr)
- [Police Employment](#police-employment)
- [Utilities](#utilities)

---

### Lookups

Functions for retrieving reference data from the CDE API.

| Function | Description |
|---|---|
| `get_agencies()` | Get all agencies from the CDE API |
| `get_offense_codes()` | Get offense codes from the CDE API |
| `get_states()` | Get state list from the CDE API |

```{r, eval=FALSE}
# List all states
get_states()

# List offense codes
get_offense_codes()

# Get all agencies (may take a moment)
agencies <- get_agencies()
```

### Summarized Crime

Functions for querying UCR Offenses Known and Clearances data.

| Function | Description |
|---|---|
| `get_agency_crime()` | Get agency-level crime data |
| `get_estimated_crime()` | Get state- or national-level estimated crime counts |
| `get_estimated_arson()` | Get estimated arson data |

```{r, eval=FALSE}
# Agency-level crime data (violent crime)
get_agency_crime("CA0010900")

# State-level estimated crime
get_estimated_crime("CA")

# National-level estimated crime
get_estimated_crime()

# Estimated arson data for a state
get_estimated_arson("CA")
```

### Arrest

Functions for querying UCR arrest statistics.

| Function | Description |
|---|---|
| `get_arrest_count()` | Get arrest offense counts |
| `get_arrest_demographics()` | Get arrestee demographics |
| `get_arrest_demographics_all()` | Get demographics for all UCR arrest offenses |
| `list_ucr_arrest_offenses()` | List all UCR arrest offense codes |

```{r, eval=FALSE}
# Total arrests for an agency
get_arrest_count(ori = "CA0010900")

# Arrest demographics for a state
get_arrest_demographics(state_abb = "CA", offense = "murder")

# All arrest offenses
list_ucr_arrest_offenses()
```

### NIBRS

Functions for querying National Incident-Based Reporting System data.

| Function | Description |
|---|---|
| `get_nibrs_victim()` | Get victim-level NIBRS data |
| `get_nibrs_offender()` | Get offender-level NIBRS data |
| `get_nibrs_offense()` | Get offense-level NIBRS data |
| `list_nibrs_offenses()` | List all NIBRS offense codes |
| `list_nibrs_victim_variables()` | List victim variable options |
| `list_nibrs_offender_variables()` | List offender variable options |
| `list_nibrs_offense_variables()` | List offense variable options |
| `list_regions()` | List all region choices |

```{r, eval=FALSE}
# Victim demographics (race)
get_nibrs_victim(ori = "CA0010900", variable = "race")

# Offender demographics
get_nibrs_offender(state_abb = "CA", offense = "murder", variable = "race")

# Offense characteristics (weapons)
get_nibrs_offense(state_abb = "CA", offense = "robbery", variable = "weapons")
```

### SHR

Functions for querying Supplemental Homicide Reports data.

| Function | Description |
|---|---|
| `get_shr()` | Get SHR homicide offense counts |

```{r, eval=FALSE}
# National-level SHR data
get_shr()

# California state-level SHR data
get_shr(state_abb = "CA")
```

### Police Employment

Functions for querying police staffing data.

| Function | Description |
|---|---|
| `get_police_employment()` | Get police staffing data |

```{r, eval=FALSE}
# Oakland PD staffing
get_police_employment("CA0010900")

# California state-level staffing
get_police_employment(state_abb = "CA")

# National-level staffing
get_police_employment()
```

### Utilities

Helper functions for working with the CDE API.

| Function | Description |
|---|---|
| `cde_base_url()` | Get the CDE API base URL |
| `cde_path()` | Build a CDE API path string |
| `cde_query()` | Build query parameters for CDE requests |
| `is_valid_ori()` | Validate an ORI code |
| `is_valid_state()` | Validate a state abbreviation |

```{r, eval=FALSE}
# Check the base URL
cde_base_url()
#> [1] "https://cde.ucr.cjis.gov/LATEST/"

# Build a path
cde_path("summarized", "national", "V")
#> [1] "summarized/national/V"

# Build query params
cde_query("01-2015", "12-2020", type = "counts")
#> $from
#> [1] "01-2015"
#> $to
#> [1] "12-2020"
#> $type
#> [1] "counts"

# Validate an ORI
is_valid_ori("CA0010900")
#> [1] TRUE

# Validate a state
is_valid_state("CA")
#> [1] TRUE
```

## Architecture

The package uses `cde_request()` as its single network seam. All API calls route through this function, which:

1. Builds the full URL from `cde_base_url()` + `cde_path()` + `cde_query()`
2. Performs the HTTP GET request via `httr::GET`
3. Parses the JSON response via `jsonlite::fromJSON()`
4. Returns a list for downstream parsing

This design makes it straightforward to mock the API in tests by injecting a custom `get_fun` argument.

## Data Sources

All data is sourced from the FBI Crime Data Explorer API at <https://cde.ucr.cjis.gov/LATEST/>. No API key is required for the current host.

## Documentation

See the full package documentation for detailed function reference and parameter descriptions.
